\section*{References}

\begin{enumerate}
\item \label{ref:ospf} OSPF Version 2 RFC 2328, John Moy, 1998.\\
\url{https://datatracker.ietf.org/doc/html/rfc2328}

\item \label{ref:isis} OSI IS-IS Intra-domain Routing Protocol, RFC 1142, 1990.\\ \url{https://datatracker.ietf.org/doc/html/rfc1142} 

\item \label{ref:ring-zone} Jadoon, R.N.; Zhou, W.; Jadoon, W.; Ahmed Khan, I. RARZ: Ring-Zone Based Routing Protocol for Wireless Sensor Networks. Appl. Sci. 2018, 8, 1023.\\ \url{https://doi.org/10.3390/app8071023}

\item \label{ref:rings} Slivkins, A.: Distance Estimation and Object Location via Rings of Neighbors, 2005.\\ \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/02/podc05_full.pdf}

\item \label{ref:sp-us} Agarwal R.; Caesar M.; Godfrey P.B.; Zhao B.Y.: Shortest Paths in Microseconds, 2013.\\ \url{https://arxiv.org/pdf/1309.0874.pdf}

\item \label{ref:social} Mensah, D.N.A.; Gao, H.; Yang, L.W. Approximation Algorithm for Shortest Path in Large Social Networks. Algorithms 2020, 13, 36.\\ \url{https://doi.org/10.3390/a13020036}

\item \label{ref:ispf} El-Sayed, Hesham \& Ahmed, Maaiz \& Jaseemuddin, Muhammad \& Petriu, Dorina. (2005). A Framework for Performance Characterization and Enhancement of the OSPF Routing Protocol.. 137-142.

\item \label{ref:fat} Wang, Ting \& Su, Zhiyang \& Xia, Yu \& Hamdi, Mounir. (2014). Rethinking the Data Center Networking: Architecture, Network Protocols, and Resource Sharing. IEEE Access. 2. 1481-1496. 10.1109/ACCESS.2014.2383439.

\end{enumerate}
